---
title: "Xing — Prognostic Score Analysis"
date: today
execute:
  echo: true
  warning: false
  message: false
---

# Error Fixed Applied

| Error | Solution |
|-------|----------|
| `day0_meld` / `day0_meld_3` use raw `scr_terli_day0` instead of capping to [1, 4] — MELD specification requires creatinine capped before calculation. The shared pipeline does this; the original code overwrites with uncapped versions. | Added `creatinine_adj_day0 = case_when(scr_terli_day0 < 1 ~ 1, scr_terli_day0 >= 4 ~ 4, TRUE ~ scr_terli_day0)` and used it in `day0_meld` and `day0_meld_3` formulas |
| `dc_meld` / `dc_meld_3` also use raw `dc_scr` without capping to [1, 4] — same MELD capping rule applies to discharge scores | Added `dc_creatinine_adj = case_when(dc_scr < 1 ~ 1, dc_scr >= 4 ~ 4, TRUE ~ dc_scr)` and used it in `dc_meld` and `dc_meld_3` formulas |
| `clif_c_score_day0` computed but never used — all models referenced `clif_c_score_new` (from shared pipeline, uses single MAP reading) instead of `clif_c_score_day0` (uses median of 4 MAP readings) | Replaced `clif_c_score_new` with `clif_c_score_day0` in all Cox, timeROC, and logistic regression models |
| `albvolume_before_terli` in `num_vars` but not in `all_vars` — silently excluded from Table 1 | Added `albvolume_before_terli` to `all_vars` |

```{r setup}
#| include: false
library(tidyverse)
library(readxl)
library(survival)
library(survminer)
library(tableone)
library(table1)
library(htmltools)
library(timeROC)
library(pROC)
library(gmodels)
library(DescTools)
library(sqldf)
library(MatchIt)
library(splines)

source(knitr::purl(
  "/Users/to909/Desktop/Terlipressin projects/Terlipressin/code/shared_pipeline.qmd",
  quiet = TRUE
))

# Convert all columns to numeric where possible (matches original pipeline)
for (col in names(master)) {
  converted <- suppressWarnings(as.numeric(master[[col]]))
  if (any(!is.na(converted))) {
    master[[col]] <- converted
  }
}
```

# CLIF-C Score Function

```{r clif-c-function}
#| code-fold: true
#| code-summary: "Show CLIF-C function"

calculate_clif_c_score <- function(tb_admit, peak_creatinine, he_grade_admission, inr_admit, map, pressor, fio2_admission, spo2_admission, age_admission, wbc_admit) {

  clif_score <- 0

  # TB Admit Score
  clif_score <- clif_score + case_when(
    is.na(tb_admit) ~ 0,
    tb_admit < 6 ~ 1,
    tb_admit >= 6 & tb_admit < 12 ~ 2,
    tb_admit >= 12 ~ 3,
    TRUE ~ 0
  )

  # Peak Creatinine Score
  clif_score <- clif_score + case_when(
    is.na(peak_creatinine) ~ 0,
    peak_creatinine < 2 ~ 1,
    peak_creatinine >= 2 & peak_creatinine < 3.5 ~ 2,
    peak_creatinine >= 3.5 ~ 3,
    TRUE ~ 0
  )

  # HE Grade Admission Score
  clif_score <- clif_score + case_when(
    is.na(he_grade_admission) ~ 0,
    he_grade_admission == 1 ~ 1,
    he_grade_admission %in% c(2, 3) ~ 2,
    he_grade_admission %in% c(4, 5) ~ 3,
    TRUE ~ 0
  )

  # INR Admit Score
  clif_score <- clif_score + case_when(
    is.na(inr_admit) ~ 0,
    inr_admit < 2 ~ 1,
    inr_admit >= 2 & inr_admit < 2.5 ~ 2,
    inr_admit >= 2.5 ~ 3,
    TRUE ~ 0
  )

  # MAP and Pressor Score
  clif_score <- clif_score + case_when(
    is.na(map) ~ 0,
    map >= 70 & pressor != 1 ~ 1,
    map < 70 & pressor != 1 ~ 2,
    !is.na(map) & pressor == 1 ~ 3,
    TRUE ~ 0
  )

  # FIO2 and SPO2 Admission Score
  if (!is.na(fio2_admission)) {
    ratio <- spo2_admission / fio2_admission
    clif_score <- clif_score + case_when(
      ratio > 357 ~ 1,
      ratio > 214 & ratio <= 357 ~ 2,
      ratio <= 214 ~ 3,
      TRUE ~ 0
    )
  }

  # Calculate CLIF-C Score
  CLIF_C_Score <- 10 * (0.33 * clif_score + 0.04 * age_admission + 0.63 * log(wbc_admit) - 2)

  return(CLIF_C_Score)
}
```


# Project-Specific Derived Variables

```{r xing-derived-variables}
#| code-fold: true
#| code-summary: "Show derived variables code"

# --- MAP day 0: median of 4 readings (zeros → NA) ---
master <- master %>%
  mutate(across(
    c(map_terli_day0_time0, map_terli_day0_time1, map_terli_day0_time2, map_terli_day0_time3),
    ~ na_if(as.numeric(.), 0)
  ))

map_terli_day0 <- master %>%
  group_by(subjectid) %>%
  summarise(
    map_terli_day0 = median(
      c(map_terli_day0_time0, map_terli_day0_time1, map_terli_day0_time2, map_terli_day0_time3),
      na.rm = TRUE
    ),
    .groups = "drop"
  )

# --- Median-impute day-0 labs for score calculation ---
master <- master %>%
  mutate(across(
    all_of(c("bili_terli_day0", "scr_terli_day0", "he_terli_day0", "alb_terli_day0",
             "inr_terli_day0", "map_terli_day0_time0", "norepi_terli_day0",
             "fio2_terli_day0", "spo2_terli_day0", "age", "wbc_terli_day0")),
    ~ ifelse(is.na(.), median(., na.rm = TRUE), .)
  ))

master <- master %>% merge(map_terli_day0, by = "subjectid", all.x = TRUE)

# --- CLIF-C score at day 0 ---
master <- master %>%
  mutate(
    clif_c_score_day0 = mapply(
      calculate_clif_c_score,
      bili_terli_day0, scr_terli_day0, he_terli_day0,
      inr_terli_day0, map_terli_day0, norepi_terli_day0,
      fio2_terli_day0, spo2_terli_day0, age, wbc_terli_day0
    )
  )

# --- Survival from terlipressin initiation ---
master <- master %>%
  mutate(dc_days_to_death = ifelse(disposition == 5, 0, dc_days_to_death)) %>%
  mutate(
    death_hosp = if_else(disposition == 5, 1L, 0L)
  ) %>%
  mutate(
    days_to_dc_init             = days_to_discharge - days_to_terli_initiation_day0,
    days_to_los_init            = days_to_lost_to_followup + days_to_discharge - days_to_terli_initiation_day0,
    days_to_lastencounter_init  = dc_days_lastencounter + days_to_discharge - days_to_terli_initiation_day0,
    days_to_after_dc_death_init = dc_days_to_death + days_to_discharge - days_to_terli_initiation_day0,

    death_status_check = case_when(
      disposition == 5 | dc_death_90days == 1 ~ 1,
      TRUE ~ 0
    ),

    time_to_death_check = case_when(
      disposition == 5                                                     ~ days_to_dc_init,
      dc_death_90days == 2                                                 ~ days_to_los_init,
      dc_death_90days == 1                                                 ~ days_to_after_dc_death_init,
      dc_death_90days == 0 & is.na(days_to_lastencounter_init)             ~ days_to_los_init,
      dc_death_90days == 0 & !is.na(days_to_lastencounter_init)            ~ days_to_lastencounter_init,
      TRUE ~ NA_real_
    ),
    death_status_90days_init = ifelse(time_to_death_check <= 90 & death_status_check == 1, 1, 0),
    time_to_death_90days_init = ifelse(time_to_death_check <= 90, time_to_death_check, 90)
  )

# --- Median-impute discharge labs ---
master <- master %>%
  mutate(
    dc_scr     = ifelse(is.na(dc_scr),     median(dc_scr,     na.rm = TRUE), dc_scr),
    dc_tbili   = ifelse(is.na(dc_tbili),   median(dc_tbili,   na.rm = TRUE), dc_tbili),
    dc_inr     = ifelse(is.na(dc_inr),     median(dc_inr,     na.rm = TRUE), dc_inr),
    dc_sna     = ifelse(is.na(dc_sna),     median(dc_sna,     na.rm = TRUE), dc_sna),
    dc_albumin = ifelse(is.na(dc_albumin), median(dc_albumin, na.rm = TRUE), dc_albumin)
  )

# --- Discharge MELD scores ---
master <- master %>%
  mutate(
    dc_creatinine_adj = case_when(dc_scr < 1 ~ 1, dc_scr >= 4 ~ 4, TRUE ~ dc_scr),
    dc_meld = round(9.57 * log(dc_creatinine_adj) + 3.78 * log(dc_tbili) + 11.20 * log(dc_inr) + 6.43),
    dc_meld_na = round(dc_meld + 1.32 * (137 - pmin(pmax(dc_sna, 125), 137)) -
                         (0.033 * dc_meld * (137 - pmin(pmax(dc_sna, 125), 137)))),
    dc_meld_na = ifelse(dc_meld_na > 11, dc_meld_na, NA),
    sex_male = as.numeric(sex_male),
    female_factor = if_else(sex_male == 0, 1.33, 0),
    dc_meld_3 = round(
      female_factor +
        4.56 * log(dc_tbili) +
        0.82 * (137 - dc_sna) -
        0.24 * (137 - dc_sna) * log(dc_tbili) +
        9.09 * log(dc_inr) +
        11.14 * log(dc_creatinine_adj) +
        1.85 * (3.5 - dc_albumin) -
        1.83 * (3.5 - dc_albumin) * log(dc_creatinine_adj) +
        6
    )
  )

# --- Day-0 MELD scores (recalculated from day-0 labs) ---
master <- master %>%
  mutate(
    creatinine_adj_day0 = case_when(scr_terli_day0 < 1 ~ 1, scr_terli_day0 >= 4 ~ 4, TRUE ~ scr_terli_day0),
    day0_meld = round(9.57 * log(creatinine_adj_day0) + 3.78 * log(bili_terli_day0) + 11.20 * log(inr_terli_day0) + 6.43),
    day0_meld_na = round(day0_meld + 1.32 * (137 - pmin(pmax(na_terli_day0, 125), 137)) -
                           (0.033 * day0_meld * (137 - pmin(pmax(na_terli_day0, 125), 137)))),
    day0_meld_na = ifelse(day0_meld_na > 11, day0_meld_na, NA),
    sex_male = as.numeric(sex_male),
    female_factor = if_else(sex_male == 0, 1.33, 0),
    day0_meld_3 = round(
      female_factor +
        4.56 * log(bili_terli_day0) +
        0.82 * (137 - na_terli_day0) -
        0.24 * (137 - na_terli_day0) * log(bili_terli_day0) +
        9.09 * log(inr_terli_day0) +
        11.14 * log(creatinine_adj_day0) +
        1.85 * (3.5 - alb_terli_day0) -
        1.83 * (3.5 - alb_terli_day0) * log(creatinine_adj_day0) +
        6
    )
  )
```


# Cox Proportional Hazards — 90-Day Mortality

## Day-0 MELD
```{r cox-day0-meld}
#| code-fold: true
cox_day0_meld <- coxph(Surv(time_to_death_90days_init, death_status_90days_init) ~ day0_meld, data = master, x = TRUE)
summary(cox_day0_meld)
```

## Day-0 MELD 3.0
```{r cox-day0-meld-3}
#| code-fold: true
cox_day0_meld_3 <- coxph(Surv(time_to_death_90days_init, death_status_90days_init) ~ day0_meld_3, data = master, x = TRUE)
summary(cox_day0_meld_3)
```

## Day-0 MELD-Na
```{r cox-day0-meld-na}
#| code-fold: true
cox_day0_meld_na <- coxph(Surv(time_to_death_90days_init, death_status_90days_init) ~ day0_meld_na, data = master, x = TRUE)
summary(cox_day0_meld_na)
```

## CLIF-C Score
```{r cox-clif-c}
#| code-fold: true
cox_clif_c_score_day0 <- coxph(Surv(time_to_death_90days_init, death_status_90days_init) ~ clif_c_score_day0, data = master, x = TRUE)
summary(cox_clif_c_score_day0)
```


# Time-Dependent ROC — 90-Day Mortality

## Day-0 MELD
```{r roc-day0-meld}
#| code-fold: true
ROC_day0_meld <- timeROC(
  T = master$time_to_death_90days_init,
  delta = master$death_status_90days_init,
  marker = master$day0_meld,
  cause = 1, weighting = "marginal",
  times = c(7, 15, 30, 60, 89), iid = TRUE
)
ROC_day0_meld
```

## Day-0 MELD 3.0
```{r roc-day0-meld-3}
#| code-fold: true
ROC_day0_meld_3 <- timeROC(
  T = master$time_to_death_90days_init,
  delta = master$death_status_90days_init,
  marker = master$day0_meld_3,
  cause = 1, weighting = "marginal",
  times = c(7, 15, 30, 60, 89), iid = TRUE
)
ROC_day0_meld_3
```

## Day-0 MELD-Na
```{r roc-day0-meld-na}
#| code-fold: true
ROC_day0_meld_na <- timeROC(
  T = master$time_to_death_90days_init,
  delta = master$death_status_90days_init,
  marker = master$day0_meld_na,
  cause = 1, weighting = "marginal",
  times = c(7, 15, 30, 60, 89), iid = TRUE
)
ROC_day0_meld_na
```

## CLIF-C Score
```{r roc-clif-c}
#| code-fold: true
ROC_clif_c_score_day0 <- timeROC(
  T = master$time_to_death_90days_init,
  delta = master$death_status_90days_init,
  marker = master$clif_c_score_day0,
  cause = 1, weighting = "marginal",
  times = c(7, 15, 30, 60, 89), iid = TRUE
)
ROC_clif_c_score_day0
```


# Pairwise AUC Comparisons (timeROC)

## MELD vs MELD 3.0
```{r compare-meld-meld3}
#| code-fold: true
timeROC::compare(ROC_day0_meld, ROC_day0_meld_3, adjusted = TRUE)
```

## MELD-Na vs MELD 3.0
```{r compare-meldna-meld3}
#| code-fold: true
timeROC::compare(ROC_day0_meld_na, ROC_day0_meld_3, adjusted = TRUE)
```

## CLIF-C vs MELD 3.0
```{r compare-clif-meld3}
#| code-fold: true
timeROC::compare(ROC_clif_c_score_day0, ROC_day0_meld_3, adjusted = TRUE)
```


# Table 1 — Stratified by 90-Day Mortality

```{r table-1}
#| code-fold: true
all_vars <- c(
  "age", "sex_male", "race", "BMI", "cirrhosis_etiology",
  "he_previous_comp", "ascites_previous_comp", "refascites_previous_comp",
  "sbp_previous_comp", "nonsbp_previous_comp", "phtn_bleed_previous_comp",
  "hcc_previous_comp", "tips_previous_comp", "aki_previous_90days",
  "mido_preadmission", "oct_before_terli", "pressors_aki",
  "admit_meld_3", "day0_meld_na", "day0_meld", "day0_meld_3",
  "baseline_scr_new", "baseline_scr", "tbili_admit", "inr_admit",
  "sodium_admit", "albumin_admit", "aclf_grade", "clif_c_score",
  "icu_admit", "criteria_1", "criteria_2", "criteria_3", "criteria_4",
  "alb_terli_discont_day", "terli_route_day0", "Discont_Day",
  "albumin_before_terli", "totalalbumin_before_terli",
  "hrs_responders", "hrs_responders_cat_2", "map_day0", "albvolume_before_terli",
  "rrt_90days", "lt_90days", "death_status_90days_init",
  "rrt_2", "location_terli_day0"
)

num_vars <- c(
  "age", "BMI", "admit_meld_3", "day0_meld_na", "day0_meld", "day0_meld_3",
  "baseline_scr_new", "baseline_scr", "tbili_admit", "inr_admit",
  "sodium_admit", "albumin_admit", "aclf_grade", "clif_c_score",
  "alb_terli_discont_day", "totalalbumin_before_terli",
  "albvolume_before_terli", "Discont_Day", "map_day0"
)

cat_vars <- setdiff(all_vars, num_vars)

create_table_one(
  all_vars = all_vars,
  cat_vars = cat_vars,
  data     = master,
  strata   = "death_status_90days_init",
  output   = "table1_death_status_90days_init.csv"
)
```


# Cross-Tabulation — MELD 3.0 vs MELD-Na Categories (Decedents)
```{r cross-tab}
#| code-fold: true
master_death <- master %>%
  filter(death_status_90days_init == 1) %>%
  mutate(
    meld_3_cat3  = case_when(day0_meld_3 < 25 ~ "<25", day0_meld_3 >= 25 & day0_meld_3 <= 35 ~ "25 - 35", day0_meld_3 > 35 ~ ">35"),
    meld_na_cat3 = case_when(day0_meld_na < 25 ~ "<25", day0_meld_na >= 25 & day0_meld_na <= 35 ~ "25 - 35", day0_meld_na > 35 ~ ">35")
  )

gmodels::CrossTable(
  master_death$meld_3_cat3,
  master_death$meld_na_cat3,
  prop.r = FALSE, prop.c = FALSE, prop.t = FALSE, prop.chisq = FALSE, chisq = FALSE
)
```


# Logistic Regression ROC — HRS Response
```{r logistic-roc}
#| code-fold: true
model_day0_meld        <- glm(hrs_responders_cat_2 ~ day0_meld,        data = master, family = "binomial")
model_day0_meld_na     <- glm(hrs_responders_cat_2 ~ day0_meld_na,     data = master, family = "binomial")
model_clif_c_score_day0 <- glm(hrs_responders_cat_2 ~ clif_c_score_day0, data = master, family = "binomial")
model_day0_meld_3      <- glm(hrs_responders_cat_2 ~ day0_meld_3,      data = master, family = "binomial")

master$pred_day0_meld    <- predict(model_day0_meld,        type = "response")
master$pred_day0_meld_na <- predict(model_day0_meld_na,     type = "response")
master$pred_day0_meld_3  <- predict(model_day0_meld_3,      type = "response")
master$pred_clif         <- predict(model_clif_c_score_day0, type = "response")

roc_meld_3 <- roc(master$hrs_responders_cat_2, master$pred_day0_meld_3, ci = TRUE, direction = "<")
roc_meld   <- roc(master$hrs_responders_cat_2, master$pred_day0_meld,   ci = TRUE, direction = "<")
roc_na     <- roc(master$hrs_responders_cat_2, master$pred_day0_meld_na, ci = TRUE, direction = "<")
roc_clif   <- roc(master$hrs_responders_cat_2, master$pred_clif,        ci = TRUE, direction = "<")
```

## MELD vs MELD 3.0
```{r delong-meld}
#| code-fold: true
roc_meld
roc.test(roc_meld, roc_meld_3, method = "delong")
```

## MELD 3.0
```{r delong-meld3}
#| code-fold: true
roc_meld_3
```

## MELD-Na vs MELD 3.0
```{r delong-meldna}
#| code-fold: true
roc_na
roc.test(roc_na, roc_meld_3, method = "delong")
```

## CLIF-C vs MELD 3.0
```{r delong-clif}
#| code-fold: true
roc_clif
roc.test(roc_clif, roc_meld_3, method = "delong")
```
